MLCommons Earthquake Science Benchmark

### 2022-02-19 paper-capstone-earthquake.tex
bio figures

\PassOptionsToPackage{hyphens}{url}
\documentclass[sigplan,screen]{acmart}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[pdf]{graphviz}
\usepackage{xpatch}
\usepackage{adjustbox}
\usepackage{makecell} 
\usepackage{ifthen}
% \usepackage[numbers]{natbib}‎

%\usepackage[colorlinks = true,
%            linkcolor = blue,
%            urlcolor  = blue,
%            citecolor = blue,
%            anchorcolor = blue]{hyperref}

@@

@@

For this study, we review the following model architectures across various hardware platforms to establish this benchmark:
\begin{itemize}
    \item Long Short-Term Memory (LSTM) Recurrent Neural Network (RNN)
    \item Spacio-Temporal Transformer Network (S2TN) as a Science Transformer
    \item Temporal Fusion Transformer (TFT)
\end{itemize}

@@

\begin{itemize}
    \item \textbf{Developer.}  The developer produces a YAML configuration file that describes the target execution details including platform, runtime dependencies, inputs, outputs, metadata about the model, and the source code.  Additionally, this configuration file offers branching paths based on what \mlcube{} runtime backend is selected (for example, the container image to use for docker can be configured differently than a kubeflow configuration).  This is so the implementation details can be explicitly addressed, if necessary, by the model developer and abstracts this complexity away from consumers.
    \item \textbf{User.}  The user of the system will receive the \mlcube{} code repository and execute a command that identifies which platform they intend to run the model on and any type of input parameters to the model detailed by the developer.
\end{itemize}

@@

\TODO{wrapfigure does somehow not work in this template}

{\bf Thomas Butler} is a Graduate Student at University of Virginia's
School of Data Science. His undergraduate degree is in Biomedical
engineering. He has over eight years of experience in the Infertility
field helping patients, jointly running a Andrology lab, and
contributing research to advance the field through joint research on
how AMH effects pregnancy outcomes and sperm antibodies effect PSA. He
has a certificate in Data Analytics from Georgia Institute of
Technology.

{\bf Robert Knuuti} is a Graduate Student at University of Virginia's
School of Data Science. He has over 10 years experience in system
architecture and software engineering, and specializes in Development
Operations and Cloud Computing. He has constructed air gapped
Continuous Integration and Continuous Delivery systems for multiple
organizations each supporting more than 100 developers and has
facilitated the construction of repeatable, tractable builds for users
of these systems.

{\bf Jake Kolessar} {is a Graduate Student at the University of
  Virginia's School of Data Science. He has a background in mechanical
  engineering and 2 years of experience as a Modeling, Simulation and
  Analysis Engineer. He has supported the software design and
  development of modeling capabilities for event simulation products
  as well as the integration of models into the simulation framework.}

{\bf Geoffrey C. Fox}
\TODO{Geoffery: fill out}

@@

{\bf Gregor von Laszewski} is a Research Professor at University of
Virginia. He has more than 30 years of experience in parallel and
distributed computing. Selected research organizations he worked in
include NASA, Argonne National Laboratory, Indiana University. He is
proud to have been contributing members of research teams that
invented hybrid parallel genetic algorithms, Grid computing, and the
first large scale academic hybrid cloud, as well as the larges
scientific bibliometric analysis of XSEDE in the world.

@@

\begin{itemize}
    %% \item add presentation as cite,
    \item Review MLCommons \cite{www-mlcommons} 
    \item Use jabref \cite{www-jabrefg-org} for citation management
    \item Frequently check github \cite{www-mlcommons-eathquake}
    \item Read up on TFT \cite{www-onnen2021}
    \item Become familiar with the Attention Paper \cite{vaswani2017attention}
    \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}
    \item Begin to learn about papermill \cite{www-papermill} \TODO{Rivanna's core configurations might be a bit limited on this front.  They do provide multiple versions of software (such as py2.7 and py3.8), but we may need to get permission to do something in userspace if we need to be very specific on a version of python.  Do we want to look into using tox, conda, or leave this up to the container ecosystem to solve?}
    \TODO{Gregor: add how to use conda and modules to switch python versions}
    \item virtualenv on rivanna for a particular python version.
    \item depends on Tensorflow
    \TODO{Below this line are objectives or targets to take the current modeling solution and mature it for other platforms / ecosystems.  Rivanna uses Lua's lmod ecosystem for jailing a process, and anaconda uses solved environments for dependencies that extend beyond just python modules.}
    \item Familiarize yourself with Rivanna's modules \cite{www-modules}  and conda \cite{www-conda} environments.
    \item pytorch \cite{www-pytorch}
    \TODO{There is interest in comparing pytorch and tensoflow}
    \item horovod \cite{www-horovod}
    \TODO{MLCommons project that can target a few platforms using a YAML contract.  Once we solve the target environment, we can likely target porting this way.}

@@

\begin{itemize}
\item Reviewed background information \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/1}}
    \begin{itemize}
    \item Earthquake Nowcasting with Deep Learning
    \item MLCommons Benchmark Presentation
    \item Attention is ALl you Need
    \item MLCommons TEvolOp Earthquake Forcasting objectives
    \item Learning on Temporal Fusion Transformers
    \item MLCommons and AI for Science illustrated by Deep Learning for Geospatial Time Series
    \end{itemize}
\item Reviewed background technologies \footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/2}}
    \begin{itemize}
    \item lmod - lua module and environment management (used on rivanna)
    \item pip / conda env - python package management
    \item SLURM - an HPC batch job runner
    \item papermill - parameterized jupyter notebooks
    \item horovod - distributed framework for running multiple deep learning frameworks
    \item \mlcube{} - MLCommons repeatable machine learning framework with multiple backends.
    \item singularity - container runtimes targeting HPC clusters
    \item docker
\end{itemize}
\item Wrote introduction to paper (in progress) \footnote{\url{https://www.overleaf.com/project/61f1e28f076f4111c3ba927e}}
    \begin{itemize}
    \item added references to references.bib in overleaf
    \end{itemize}
\item Revised Intro to Python guide for conda installations so it doesn't interfere with pyenv or other shell commands.\footnote{\url{https://github.com/cloudmesh-community/book/pull/574}} \footnote{\url{https://github.com/cloudmesh-community/book/pull/575}}
\item Solved conda environment and requirements.txt for colab notebook
\footnote{\url{https://github.com/Data-ScienceHub/mlcommons-science/issues/12}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/1}} \footnote{\url{https://github.com/laszewsk/mlcommons/pull/3}}
\item build docker container image that can run python 3.9.7 and conda
\item verified access to GPU allocations in Rivanna
\end{itemize}

>>>>

\PassOptionsToPackage{hyphens}{url}
\documentclass[sigplan,screen]{acmart}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[pdf]{graphviz}
\usepackage{xpatch}
\usepackage{adjustbox}
\usepackage{makecell} 
\usepackage{ifthen}

@@

\newcommand{\GITISSUE}[1]{\href{https://github.com/Data-ScienceHub/mlcommons-science/issues/#1}{GitHub Issue #1}}}
\newcommand{\GITPULL}[1]{\href{https://github.com/laszewsk/mlcommons/pull/#1}{GitHub Pull #1}}}

@@

For this study, we review the following model architectures across
various hardware platforms to establish this benchmark:

\begin{itemize}
    \item Long Short-Term Memory (LSTM) Recurrent Neural Network (RNN)
    \item Spacio-Temporal Transformer Network (S2TN) as a Science Transformer
    \item Temporal Fusion Transformer (TFT)
\end{itemize}

@@

\begin{itemize}

    \item \textbf{Developer.}  The developer produces a YAML
      configuration file that describes the target execution details
      including platform, runtime dependencies, inputs, outputs,
      metadata about the model, and the source code.  Additionally,
      this configuration file offers branching paths based on what
      \mlcube{} runtime backend is selected (for example, the
      container image to use for docker can be configured differently
      than a kubeflow configuration).  This is so the implementation
      details can be explicitly addressed, if necessary, by the model
      developer and abstracts this complexity away from consumers.

    \item \textbf{User.}  The user of the system will receive the
      \mlcube{} code repository and execute a command that identifies
      which platform they intend to run the model on and any type of
      input parameters to the model detailed by the developer.
\end{itemize}

@@

\newcommand{\BIOFIG}[3]{%

  {\bf #2} #3
  %\begin{wrapfigure}[l]{.2\columnwidth}
  \includegraphics[width=.2\columnwidth]{#1}%
  %\end{wrapfigure}
}

\BIOFIG{images/bio/gregor.png}{Thomas Butler}{is a Graduate Student at
  University of Virginia's School of Data Science. His undergraduate
  degree is in Biomedical engineering. He has over eight years of
  experience in the Infertility field helping patients, jointly
  running a Andrology lab, and contributing research to advance the
  field through joint research on how AMH effects pregnancy outcomes
  and sperm antibodies effect PSA. He has a certificate in Data
  Analytics from Georgia Institute of Technology.}

\BIOFIG{images/bio/gregor.png}{Robert Knuuti}{is a Graduate Student at
  University of Virginia's School of Data Science. He has over 10
  years experience in system architecture and software engineering,
  and specializes in Development Operations and Cloud Computing. He
  has constructed air gapped Continuous Integration and Continuous
  Delivery systems for multiple organizations each supporting more
  than 100 developers and has facilitated the construction of
  repeatable, tractable builds for users of these systems.}

\BIOFIG{images/bio/gregor.png}{Jake Kolessar}{is a Graduate Student at
  the University of Virginia's School of Data Science. He has a
  background in mechanical engineering and 2 years of experience as a
  Modeling, Simulation and Analysis Engineer. He has supported the
  software design and development of modeling capabilities for event
  simulation products as well as the integration of models into the
  simulation framework.}

\BIOFIG{images/bio/gregor.png}{Geoffrey C. Fox}{TBD TBD TBD TBD TBD
  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
  TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD TBD
  TBD TBD TBD TBD TBD TBD TBD TBD}

@@

\BIOFIG{images/bio/gregor.png}{Gregor von Laszewski}{is a Research
  Professor at University of Virginia.  He has more than 30 years of
  experience in parallel and distributed computing.  Selected research
  organizations he worked at include NASA, Argonne National
  Laboratory, Indiana University. He is proud to have been
  contributing members of research teams that invented hybrid parallel
  genetic algorithms, Grid computing, and the first large scale
  academic hybrid cloud, as well as the larges scientific bibliometric
  analysis of XSEDE in the world.}

@@

\begin{itemize}
    %% \item add presentation as cite,
    \item Review MLCommons \cite{www-mlcommons} 
    \item Use jabref \cite{www-jabrefg-org} for citation management
    \item Frequently check github \cite{www-mlcommons-eathquake}
    \item Read up on TFT \cite{www-onnen2021}
    \item Become familiar with the Attention Paper \cite{vaswani2017attention}
    \item Learn how to selfdeploy jupyterlab \cite{www-jupyterlab}

    \item Begin to learn about papermill \cite{www-papermill}
      \TODO{Rivanna's core configurations might be a bit limited on
        this front.  They do provide multiple versions of software
        (such as py2.7 and py3.8), but we may need to get permission
        to do something in userspace if we need to be very specific on
        a version of python.  Do we want to look into using tox,
        conda, or leave this up to the container ecosystem to solve?}

      \TODO{Gregor: add how to use conda and modules to switch python versions}
    \item virtualenv on rivanna for a particular python version.
    \item depends on Tensorflow

    \TODO{Below this line are objectives or targets to take the
      current modeling solution and mature it for other platforms /
      ecosystems.  Rivanna uses Lua's lmod ecosystem for jailing a
      process, and anaconda uses solved environments for dependencies
      that extend beyond just python modules.}

  \item Familiarize yourself with Rivanna's modules \cite{www-modules}  and conda \cite{www-conda} environments.
    \item pytorch \cite{www-pytorch}
    \TODO{There is interest in comparing pytorch and tensoflow}
    \item horovod \cite{www-horovod}

      \TODO{MLCommons project that can target a few platforms using a
        YAML contract.  Once we solve the target environment, we can
        likely target porting this way.}

@@

\begin{itemize}
\item Reviewed background information \GITISSUE{1}
    \begin{itemize}
    \item Earthquake Nowcasting with Deep Learning
    \item MLCommons Benchmark Presentation
    \item Attention is ALl you Need
    \item MLCommons TEvolOp Earthquake Forcasting objectives
    \item Learning on Temporal Fusion Transformers
    \item MLCommons and AI for Science illustrated by Deep Learning for Geospatial Time Series
    \end{itemize}
\item Reviewed background technologies \GITISSUE{2}
    \begin{itemize}
    \item lmod - lua module and environment management (used on rivanna)
    \item pip / conda env - python package management
    \item SLURM - an HPC batch job runner
    \item papermill - parameterized jupyter notebooks
    \item horovod - distributed framework for running multiple deep learning frameworks
    \item \mlcube{} - MLCommons repeatable machine learning framework with multiple backends.
    \item singularity - container runtimes targeting HPC clusters
    \item docker
\end{itemize}
\item Wrote introduction to paper (in progress) \footnote{\url{https://www.overleaf.com/project/61f1e28f076f4111c3ba927e}}
    \begin{itemize}
    \item added references to references.bib in overleaf
    \end{itemize}

  \item Revised Intro to Python guide for conda installations so it
    doesn't interfere with pyenv or other shell
    commands.\footnote{\url{https://github.com/cloudmesh-community/book/pull/574}} \footnote{\url{https://github.com/cloudmesh-community/book/pull/575}}

  \item Solved conda environment and requirements.txt for colab notebook

    \GITISSUE{12},
    \GITPULL{1},
    \GITPULL{3}.

\item build docker container image that can run python 3.9.7 and conda
\item verified access to GPU allocations in Rivanna

\end{itemize}
